Setup

### 2025-04-30 setup.qmd
update setup

R and Python scripts are in the `R/` and `python/` directories respectively.

@@

* H3 hexagons at resolution `R = 2`: 
  - Average Area (km2): `86801.78`
  - Average edge length (Km): `182.51`
* H3 hexagons at resolution `R = 3`: 
  - Average Area (km2): `12393.43`
  - Average edge length (Km): `68.98`
* bounding box approximating EUROCONTROL Area (decimal degrees):
  - xmin = -25.488281
  - ymin = 26.638253
  - xmax = 45.407181
  - ymax = 71.864780
* bounding box containing the HEX at resolution 2 that match
  the bounding box approximating EUROCONTROL Area:
  - xmin = -27.78166
  - xmax =  48.56037
  - ymin =  24.90098
  - ymax =  72.85065

@@

```{r cells-map}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/figures/cells_2.png")
```

H3 **[cells]{colour="#ff0000"}** ([red]{colour="#ffffff" bg-colour="#ff0000"})
at resolution 2 and
their **[centers]{colour="#0000ff"}** ([blue]{colour="#ffffff" bg-colour="#0000ff"}) covering EUROCONTROL Area.

@@

This is achieved via script `export_data.R`.
The relevant trajectories are saved in
`data-raw/trjs_{D}.parquet`, i.e. `data-raw/trjs_2024-08-01.parquet`.

### Trajectory Resampling
We load the trajectories and resample them at 30s using the Python `traffic` library.
This is performed by the `resample_aug.py` and `resample_dec.py`
scripts and is time consuming (40 min each).
Execute as:

> $ python resample_aug.py

The result is in (52min 51s)
`data/trajectories_2024-08-01_resampled_30s.parquet`

### Fast H3 Labelling
The script `prepare_pre_density.R` uses DuckDB for fast labelling
of each lon/lat to the relevant H3 cell (at resolution `R`)
and filtering for samples between bbox.
Keep only point in the hex cells of the bbox and for day `D`.
The result is saved in `trajectories_<YYYY-MM-DD>_resampled_30s_bbox_res_3.parquet` for date `D`.

@@

The script `prepare_density.R` calculates the occupancy per hour `H` of `D` 
for each cell at resolution `R`
.

>>>>

R scripts are in the `R/` directory.

You need duckdb 1.2.2 with `icu` extension and `h3` community extension
installed (extension location hardcoded in `aviodebris::hexagonize`).

@@

* H3 hexagons at resolution `R = 3`: 
  - Average Area (km2): `12393.43`
  - Average edge length (Km): `68.98`
* bounding box approximating EUROCONTROL Area (decimal degrees) in the
  `aviodebris` packages (enclosing FIRs of Member States)

@@

```{r cells-map}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/figures/cells_3.png")
```

H3 **[cells]{colour="#ff0000"}** ([red]{colour="#ffffff" bg-colour="#ff0000"})
at resolution 3 and
their **[centers]{colour="#0000ff"}** ([blue]{colour="#ffffff" bg-colour="#0000ff"}) covering EUROCONTROL Area (([blue]{colour="#ffffff" bg-colour="#0000ff"}) polygon).

@@

This is achieved via script `export_data.R`.
The relevant trajectories are saved in `data-raw/trjs/trjs_{D}.parquet`,
i.e. `data-raw/trjs/trjs_2024-08-01.parquet`.

### Trajectory Resampling
We load the trajectories and resample them at 30s using helper function in the
`aviodebris` package, 10min per day.  
The result is in 
`data/trajectories_2024-08-01_resampled_30s.parquet`

### Fast H3 Labelling
The `hexagonize` function in `aviodebris` uses DuckDB (plus ICU extension
and H3 community extension) for fast labelling
of each lon/lat to the relevant H3 cell (at resolution `R`)
and filtering for samples within EUROCONTROL bbox.
Keep only point in the hex cells of the bbox and for day `D`.
The result is saved in `trajectories_<YYYY-MM-DD>_resampled_30s_bbox_res_<R>.parquet` for date `D`.

@@

The function `traffic_density_hourly` in the package `aviodebris` calculates
the occupancy per hour `H` of `D` for each cell at resolution `R`.
A file for each day is saved as
`traffic_density_<YYYY-MM-DD>_res_<R>_hourly.parquet` for date `D`.
The results are summarised per aircraft type.

## Collision and Casuality Risk
The function `collision_and_casuality_risk_expectation_hourly`
calculates both collision and casuality risk on a hourly basis.
